---
title: "Stat130 Project 2"
author: "Xiaohui Ma"
date: "2024-11-17"
output:
  html_document: default
  pdf_document: default
---

# Process the data

```{r}
#import the data
p2_data <- read.csv("/Users/maxiaohui/Desktop/Stat130 project 2/sunroof.csv")

#clean the data
cleaned_data <- p2_data[p2_data$state_name != "NULL", ] # remove the row that contain the NULL value in state_name

#summary of the data
dim(cleaned_data)
str(cleaned_data)

```

# Simple Random Sampling

### Decide the sample size

```{r}
size_srs<-round(909 * 0.35)  #cleaned_data have 909 row and we sample 35% data of the population 
size_srs
```

### Create an sample for SRS

```{r}
set.seed(1)
inds = sample(1:nrow(cleaned_data), size_srs, replace = FALSE) 
sample_srs = cleaned_data[inds, ]

```

### Calculation

```{r}
# estimating the total number of carbon offset by using srs
N_srs = nrow(cleaned_data)  # population size
y_bar = mean(sample_srs$carbon_offset_metric_tons) # sample mean
total_srs <- N_srs * y_bar
total_srs

# calculate the standard error of estimator
n_srs = 318    # sample size
s2 <- (1/(n_srs-1))* sum((sample_srs$carbon_offset_metric_tons - y_bar)^2)  #sample variance 
var_srs <- (N_srs^2)* (s2/n_srs)*(1-(n_srs/N_srs))   # variance of estimator
se_srs <- sqrt(var_srs)
se_srs

#calculate the 95% confidence interval of real t
lower_srs <- total_srs - qt(0.975, n_srs-1) * se_srs
upper_srs <- total_srs + qt(0.975, n_srs-1) * se_srs
CI_srs <- c(lower_srs, upper_srs)
CI_srs

```
# Stratified sampling

#### (separate the whole population into different stratums h base on states, then randomly sample several data from each stratums)

### Decide the sample size by using proportional allocation

```{r}
Nh = table(cleaned_data$state_name)  #number of data in each stratum h
state_proportions <- Nh/sum(Nh)
nh = round(state_proportions * n_srs)  #number of data sampled from each stratum h
n_str = sum(nh)  #total sample size
```

### Create an sample for Stratified sampling

```{r}
library(dplyr)

set.seed(2)
sample_str <- cleaned_data %>% 
  group_by(state_name) %>%
  group_modify(~ slice_sample(.x, n = nh[as.character(.y$state_name)]))
```

### summarize the sampled data

```{r}
summary_str <- sample_str %>%
  group_by(state_name) %>%
  summarise(
    mean_carbon_offset = mean(carbon_offset_metric_tons),
    total_carbon_offset = sum(carbon_offset_metric_tons),
    nh = n()
  )

new_Nh <- Nh[names(Nh) != "Vermont"] # remove the state "Vermont" since sampled data doesn't contain it

#calculate the population variance for each stratum. (We don't use sample variance here since some sample number in each strutum is only one)
pop_sd <-cleaned_data %>% 
  group_by(state_name) %>%
  summarise(sd_value = sd(carbon_offset_metric_tons)) %>%
  filter(!is.na(sd_value))

# add Nh in the summary table
summary_str$Nh <- new_Nh 

# add sd of each strutum in the summary table
summary_str$sd_carbon_offset <- pop_sd$sd_value  

summary_str

#Summary the Carbon offset of each Strata
summary(summary_str$total_carbon_offset)

```

### Calculation

```{r}
# estimating the total number of carbon offset by using str
total_str <- sum(summary_str$Nh * summary_str$mean_carbon_offset)
total_str

# calculate the standard error of estimator
var_str <- sum((1-(summary_str$nh/summary_str$Nh))* (summary_str$Nh^2) * ((summary_str$sd_carbon_offset^2)/summary_str$nh)) #variance
se_str = sqrt(var_str)
se_str

#calculate the 95% confidence interval of real t
H = length(unique(sample_str$state_name))  #the number of strutum
lower_str <- total_str - qt(0.975, n_str-H) * se_str
upper_str <- total_str + qt(0.975, n_str-H) * se_str
CI_str <- c(lower_str, upper_str)
CI_str
```

### Create a side-by-side boxplot

```{r}
boxplot(carbon_offset_metric_tons ~ state_name, 
        data = sample_str, 
        col = rainbow(length(unique(sample_str$state_name))), 
        main = "Carbon offset across states",
        xlab = "States",
        ylab = "Carbon offest (in tons)", 
        las = 2,                
        cex.axis = 0.8,
        cex.lab = 1.2, 
        notch = FALSE,           
        outline = FALSE )
```

# One-stage cluster sampling

#### (separate the whole population into different cluster base on states. First, randomly pick several cluster and then sample all value in each selected cluster.)

### Decide the sample size

```{r}
N_1c = length(unique(cleaned_data$state_name))      #total number of clusters in the population
n_1c = 17         #total number of sampled clusters
```

### Create an sample for One-stage cluster sampling

```{r}
states = labels(table(cleaned_data$state_name))[[1]] # 50 clusters
set.seed(3)
states_sample = sample(states, size = 17, replace = F)
sample1_clus = cleaned_data[cleaned_data$state_name %in% states_sample, ]

Mi_1c = table(sample1_clus$state_name) # Mi = mi = the number of all data in each sampled cluster
```

### summarize the sampled data

```{r}
summary_1clus <- sample1_clus %>%
  group_by(state_name) %>%
  summarise(
    total_carbon_offset = sum(carbon_offset_metric_tons),
    mean_carbon_offset = mean(carbon_offset_metric_tons)
  )

summary_1clus$Mi <- Mi_1c
summary_1clus

#Summary the Carbon offset of each cluster
summary(summary_1clus$total_carbon_offset)
```

### Calculation

```{r}
# estimating the total number of carbon offset by using one-stage cluster sampling method
total_1clus <-(N_1c/n_1c) * sum(summary_1clus$total_carbon_offset)
total_1clus

# calculate the standard error of estimator
st2_1clus <- (1/(n_1c - 1)) * sum((summary_1clus$total_carbon_offset - total_1clus)^2)
var_1clus = (N_1c)^2 * (1-(n_1c/N_1c)) * (st2_1clus/n_1c)
se_1clus = sqrt(var_1clus)
se_1clus

#calculate the 95% confidence interval of real t
lower_1clus <- total_1clus - qt(0.975, n_1c - 1) * se_1clus
upper_1clus <- total_1clus + qt(0.975, n_1c - 1) * se_1clus
CI_1clus <- c(lower_1clus, upper_1clus)
CI_1clus

```

### Create a side-by-side boxplot

```{r}
boxplot(carbon_offset_metric_tons ~ state_name, 
        data = sample1_clus, 
        col = rainbow(length(unique(sample1_clus$state_name))), 
        main = "Carbon offset across states",
        xlab = "States",
        ylab = "Carbon offest (in tons)", 
        las = 2,                
        cex.axis = 0.8,
        cex.lab = 1.2, 
        notch = FALSE,           
        outline = FALSE )
```

# Two-stage cluster sample

#### （separate the whole population into different cluster base on states. First, randomly pick several cluster and then randomly sample several data from the selected cluster again.）

### Decide the sample size

```{r}
N_2c = length(unique(cleaned_data$state_name))       #total number of clusters in the population
n_2c = 27        #total number of sampled clusters
```

### Create an sample for two-stage cluster sampling

```{r}
states = labels(table(cleaned_data$state_name))[[1]] # 50 clusters
set.seed(4)
states_sample2 = sample(states, size = 27, replace = F)  
sampled_cluster = cleaned_data[cleaned_data$state_name %in% states_sample2, ]


sample2_clus <- sampled_cluster %>%
  group_by(state_name) %>%
  sample_frac(size = 0.65)  #sample 65% of data in each sampled cluster
sample2_clus

Mi_2c = table(sampled_cluster$state_name) # the number of all data in each sampled cluster

mi <- sample2_clus %>%    # number of data sampled from each Mi
  group_by(state_name) %>%
  summarise(mi = n())
```

### summarize the sampled data

```{r}
summary_2clus <- sample2_clus %>%
  group_by(state_name) %>%
  summarise(
    total_carbon_offset = sum(carbon_offset_metric_tons),
    mean_carbon_offset = mean(carbon_offset_metric_tons), 
    si2 = var(carbon_offset_metric_tons)
  )

summary_2clus$Mi <- Mi_2c
summary_2clus$mi <- mi$mi

  
summary_2clus <- na.omit(summary_2clus) #remove the row contain NA
summary_2clus



#Summary the Carbon offset of each cluster
summary(summary_2clus$total_carbon_offset)

```

### Calculation

```{r}
# estimating the total number of carbon offset by using two-stage cluster sampling method
total_2clus = (N_2c/n_2c) * sum(summary_2clus$Mi * summary_2clus$mean_carbon_offset)
total_2clus

# calculate the standard error of estimator
st2_2clus = (1/(n_2c-1)) * sum((summary_2clus$total_carbon_offset - (total_2clus/N_2c))^2)
var_2clus = (N_2c^2) * (1-(n_2c/N_2c)) * (st2_2clus/n_2c) + (N_2c/n_2c) * sum((1-(summary_2clus$mi/summary_2clus$Mi)) * (summary_2clus$Mi^2) *(summary_2clus$si2)/summary_2clus$mi)
se_2clus = sqrt(var_2clus)
se_2clus

#calculate the 95% confidence interval of real t
lower_2clus <- total_2clus - qt(0.975, n_2c - 1) * se_2clus
upper_2clus <- total_2clus + qt(0.975, n_2c - 1) * se_2clus
CI_2clus <- c(lower_2clus, upper_2clus)
CI_2clus


```

### Create a side-by-side boxplot

```{r}
boxplot(carbon_offset_metric_tons ~ state_name, 
        data = sample2_clus, 
        col = rainbow(length(unique(sample2_clus$state_name))), 
        main = "Carbon offset across states",
        xlab = "States",
        ylab = "Carbon offest (in tons)", 
        las = 2,                
        cex.axis = 0.8,
        cex.lab = 1.2, 
        notch = FALSE,           
        outline = FALSE )
```
